\section{Flow analysis}
Several models can be used to describe an application and represent its code.  
Different models may emphasize different aspects of program behavior — such as \textbf{control flow} or \textbf{data dependence}. 
\\\textbf{Flow analysis} focuses on understanding how information moves through the program.  
By analyzing both control and data dependencies, it helps identify relationships among variables, track how data values are produced and consumed, and evaluate the effects of data changes across the codebase.  
\\\\Program analysis and test design techniques often leverage data flow information to enhance their effectiveness.  

\subsection{Control-flow graph}
A \textbf{Control-Flow Graph (CFG)} is defined as follows:
\[
CFG = (N, E, n_e, n_x)
\]
\begin{itemize}
    \item $N$: \textbf{set of nodes}, one for each \textbf{statement} (instruction) of the program;
    \item $E \subseteq N \times N$: \textbf{set of edges}, where $(n, m) \in E$ if statement $m$ can be executed immediately after $n$;
    \item $n_e$: \textbf{entry node} of the program;
    \item $n_x$: \textbf{exit node} of the program.
\end{itemize}

\paragraph{Path:} represents a possible program execution. It is composed of a sequence of nodes and edges starting from the entry node and ending at a terminal node.

\paragraph{Linearly independent path:} a path is \textbf{linearly independent} if it introduces at least one new edge that has not been traversed by any previously defined path.  
These paths are used to identify distinct behaviors implemented in a program. Typically, a test case should be defined to verify each independent path.
\subsubsection{Def-Use pairs}
A \textbf{def-use pair} associates a point in a program code where a value is defined with a point where it's used.
\paragraph{Definition:} where a variable gets a value:
\begin{itemize}
    \item Declaration;
    \item Initialization;
    \item Assignment;
    \item Values received by a parameter.
\end{itemize}
\paragraph{Use:} extraction of a value from a variable:
\begin{itemize}
    \item Expressions;
    \item Conditional statements;
    \item Parameter passing;
    \item returns.
\end{itemize}
\paragraph{Def-clear path:} a path along the CFG from a definition to a use of the same variable, without another definition of the variable in between.

\subsubsection{Data Dependence Graph}
A direct data dependence graph is defined using the two definitions  we just gave:
\begin{itemize}
    \item Its nodes are the same as in the CFG;
    \item Its edges are the def-use pairs.
\end{itemize}
This graph can be used to identify \textbf{data dependence}, where values are from.
\subsubsection{Control Dependence Graph}
Also a \textbf{control dependence graph} can be defined as follows:
\begin{itemize}
    \item Its nodes are the same as the same as in the CFG;
    \item Its edges are unlabeled, direct control dependencies.
\end{itemize}
This type of graph shows \textbf{control dependence}: which \textbf{statement controls} whether a statement is executed or not. 
\subsection{Data-Flow analysis}
\subsubsection{Reaching definition}
There is an association $(d,u)$ between a \textbf{definition of a variable} \verb|v| at code statement $d$ and a \textbf{use of variable} \verb|v| at code statement $u$ iff:
\begin{itemize}
    \item there is at least one control flow path from $d$ to $u$;
    \item there is no intervening definition of \verb|v|.
\end{itemize}
In this case, we say that $v_d$ \textbf{reaches} $v_u$.

\paragraph{Normal graph exploration}
Even if we consider a loop-free path, the number of paths can be exponentially larger than the number of nodes and edges. Practical algorithms therefore do not search every individual path. Instead, they summarize the reaching definitions at a node over all the paths reaching that node.

\paragraph{DF Algorithm:} an efficient algorithm for \textbf{computing reaching definitions}, and other properties, is based on the way \textbf{reaching definitions} at one node are related to the one of its adjacent nodes.
\\\\Suppose we are calculating the \textbf{reaching definitions} of node $n$, and there is an edge $(p,n)$ from an immediate predecessor, node $p$:
\begin{itemize}
    \item if the predecessor node $p$, assigns a value to a variable \verb|v|, then the definition $v_p$ \textbf{reaches} $n$;
    \item If a definition $v_p$ of a variable \verb|v| reaches a predecessor node $p$, then the definition is propagated on from p to n. 
\end{itemize}
\paragraph{Formal definition:} we can formally define the idea expressed by the algorithm we just cited. At each node $n$, we have:
\begin{itemize}
    \item \textbf{Reaching definitions} $(In(n))$: definitions flowing out of th predecessor nodes $m$. Can be defined as
    \[\bigcup_{m\in pred(n)} Out(m)\]
    \item \textbf{Out flow} $Out(n)$: the output of a node n can formally be defined as:
    \[Out(n) = In(n) \setminus Kill(n) \cup Gen(n) \]
    Where:
    \begin{itemize}
        \item $Gen(n) = \{v_n| v\text{ is defined or modified at }n\}$
        \item $Kill(n) = \{v_x| v \text{ is defined or modified at x AND } v \text{ is modified at} n\}$
    \end{itemize}
\end{itemize}
\subsubsection{Meet over path}
Another solution for data-flow analysis is the \textbf{Meet Over Path} (MOP).
\\At a node $n$, the MOP is defined as:
\[MOP[n] = \bigwedge_{p\in P_{n}} f_p(T)\]
where:
\begin{itemize}
    \item $P_n$ the set of all path from the entry node to node n;
    \item $f_p$ is the composition of transfer function along path $p$;
    \item $T$ is the initial information
    \item $\wedge$ represents the meet operator.
\end{itemize}
\paragraph{Exact solution:} the exact solution requires, instead of considering all path from the entry node, to the node n, to just consider the \textbf{feasible paths}.
\[EX[n] = \bigwedge_{p\in feas(P_{n})} f_p(T)\]
This is an \textbf{undecidable} problem.